Persistence in functional topology and a correction to a theorem of Morse

### 2021-07-26 connectivity.tex
Improve code

For the rest of this section, we will assume that $\H$ is a homology theory that has either the open or the compact Mayer-Vietoris property and for which there is $n_0$ such that $\H_{n}$ is 0 for all $n \leq n_0$.
Note that this includes singular homology with field coefficients, which, like any homology theory in the sense of the Eilenberg-Steenrod axioms \cite[Section I]{MR0050886}, has the open Mayer-Vietoris property, and it also includes \v{C}ech homology with field coefficients, which has the compact Mayer-Vietoris property.

\begin{defi} \label{defi:local_connectedness}
	For $n \in \Z$ a continuous map is said to be \textit{$n$-homologically small} or $\HS_n$ if the image of the map induced by $\H_{n}$ is finite dimensional.
	We omit references to $n$ if the condition holds for all integers.
\end{defi}

\begin{defi} \label{defi:local_connectedness_filtrations}
	The sublevel set filtration of a function $f \colon X \to \R$ is said to be \emph{locally homologically small} or $\HLC$ if for any $x \in X$, any neighborhood $V$ of $x$, and any index $t > f(x)$, there is an index $s$ with  $f(x) < s < t$ and a neighborhood $U$ of $x$ with $U \subseteq V$ such that the inclusion $f_{\leq s} \cap U \to f_{\leq t} \cap V$ is $\HS$.
\end{defi}
Similarly, we have the following stronger version of this property.
\begin{defi}
	The sublevel set filtration of a function $f \colon X \to \R$ is called \emph{strongly locally homologically small} or \emph{strongly $\HLC$} if for any $x \in X$, any neighborhood $V$ of $x$, and any pair of indices $s,t$ with $f(x) < s < t$ there is a neighborhood $U$ of $x$ with $U \subseteq V$ such that the inclusion $f_{\leq s} \cap U \to f_{\leq t} \cap V$ is $\HS$.
\end{defi}

@@

\begin{thm}\label{thm:hlc_to_strong_hlc}
    If the sublevel set filtration of a continuous function $f \colon X \to \R$ is $\HLC$, then it is also strongly $\HLC$.
\end{thm}
\begin{proof}
    Fix $x \in X$, a neighborhood $V$ of $x$ and indices $f(x) < s < t$.
	We need to show that there is a neighborhood $U \subseteq V$ of $x$ such that the inclusion $f_{\leq s} \cap U \to f_{\leq t} \cap V$ is $\HS$.

@@

\begin{lem}\label{l: restriction of LHS filtration}
    Let $f \colon X \to \R$ be a function whose sublevel set filtration is strongly $\HLC$.
    Fix $u \in \R$ and let $g \colon Y \to \R$ be the restriction of $f$ to the sublevel set $Y = f_{\leq u}$.
    Then, the sublevel set filtration defined by $g$ is strongly $\HLC$ below $u$.
\end{lem}

@@

    We define $\Sigma_{s, t}$ to be the collection of all open subsets $V \subseteq X$ whose closure $\overline{V}$ is compact, contained in $L$, and has an open neighborhood $U$ with 
	$\overline{V} \subseteq U \subseteq L$
	for which there exists $s' \in (s,\, t)$ such that the inclusion
    $U \cap f_{\leq s'} \to L \cap f_{\leq t}$
	is $\HS_n$.
	We will show that $\Sigma_{s, t}$ has the following three properties:
	\begin{enumerate}
	    \item Any point $x \in L \cap f_{\leq s}$ has a neighborhood $V_x \in \Sigma_{s,t}$.
	    \item If $V_1,\, V_2 \in \Sigma_{s,t}$, then $V_1 \cup V_2 \in \Sigma_{s,t}$.
	    \item For each $V \in \Sigma_{s,t}$ the inclusion 
	    $V \cap f_{\leq s} \to L \cap f_{\leq t}$ 
	    is $\HS_n$.
	\end{enumerate}

@@

\begin{proof}[Proof of \cref{t:strong local connectedness implies q-tameness}]
    By definition, the sublevel set filtration of $f$ is q-tame if and only if the inclusion $f_{\leq s} \to f_{\leq t}$ is $\HS$ for all pairs $s < t$. 
    Choose $u \in \R$ with $u > t$ and let $g \colon Y \to \R$ be the restriction of $f$ to the sublevel set $Y=f_{\leq u}$. 
    Since we assume $f$ to induce a strongly $\HLC$ sublevel set filtration, by \cref{l: restriction of LHS filtration} the sublevel set filtration of $g$ is strongly $\HLC$ below $u$. 
    Clearly, the sublevel set filtration of $g$ is also compact, and its domain $Y$ is locally compact being a compact Hausdorff space by assumption.
    Thus, we can apply \cref{l:key lemma for q-tameness} to the filtration $g_{\leq \bullet}$ with $C = f_{\leq s}$ and $L = Y$ to obtain that the inclusion $f_{\leq s} = C \cap g_{\leq s} \to L \cap g_{\leq t} = f_{\leq t}$ is $\HS$.
\end{proof}

Combining \cref{thm:hlc_to_strong_hlc} with \cref{t:strong local connectedness implies q-tameness} also yields the following result for continuous functions.

\begin{cor}\label{c: q-tameness for continuous functions}
    If the sublevel set filtration of a continuous function $f \colon X \to \R$ is compact and $\HLC$, then it is also q-tame.
\end{cor}

>>>>

For the rest of this section, we will assume that $\H$ is a homology theory that has either the open or the compact Mayer-Vietoris property and for which there is $n_0$ such that $\H_{n}$ is 0 for all $n \leq n_0$.
Note that this includes singular homology with field coefficients, which, like any homology theory in the sense of the Eilenberg-Steenrod axioms \cite[Section I]{Eilenberg.1952}, has the open Mayer-Vietoris property, and it also includes \v{C}ech homology with field coefficients, which has the compact Mayer-Vietoris property.

\begin{defi}
	For $n \in \Z$ a continuous map is said to be \emph{$n$-homologically small} or $\HS_n$ if the image of the map induced by $\H_{n}$ is finite dimensional.
	We omit references to $n$ if the condition holds for all integers.
\end{defi}

\begin{defi}
	The sublevel set filtration of a function $f \colon X \to \R$ is said to be \emph{locally homologically small} or $\HLC$ if for any $x \in X$, any neighborhood $V$ of $x$, and any index $t > f(x)$, there is an index $s$ with $f(x) < s < t$ and a neighborhood $U$ of $x$ with $U \subseteq V$ such that the inclusion $f_{\leq s} \cap U \to f_{\leq t} \cap V$ is $\HS$.
\end{defi}

Similarly, we have the following stronger version of this property.

\begin{defi}
	The sublevel set filtration of a function $f \colon X \to \R$ is called \emph{strongly locally homologically small} or \emph{strongly $\HLC$} if for any $x \in X$, any neighborhood $V$ of $x$, and any pair of indices $s,t$ with $f(x) < s < t$ there is a neighborhood $U$ of $x$ with $U \subseteq V$ such that the inclusion $f_{\leq s} \cap U \to f_{\leq t} \cap V$ is $\HS$.
\end{defi}

@@

\begin{thm} \label{t:hlc to strong hlc}
	If the sublevel set filtration of a continuous function $f \colon X \to \R$ is $\HLC$, then it is also strongly $\HLC$.
\end{thm}

\begin{proof}
	Fix $x \in X$, a neighborhood $V$ of $x$ and indices $f(x) < s < t$.
	We need to show that there is a neighborhood $U \subseteq V$ of $x$ such that the inclusion $f_{\leq s} \cap U \to f_{\leq t} \cap V$ is $\HS$.

@@

\begin{lem} \label{l:restriction of LHS filtration}
	Let $f \colon X \to \R$ be a function whose sublevel set filtration is strongly $\HLC$.
	Fix $u \in \R$ and let $g \colon Y \to \R$ be the restriction of $f$ to the sublevel set $Y = f_{\leq u}$.
	Then, the sublevel set filtration defined by $g$ is strongly $\HLC$ below $u$.
\end{lem}

@@

	We define $\Sigma_{s, t}$ to be the collection of all open subsets $V \subseteq X$ whose closure $\overline{V}$ is compact, contained in $L$, and has an open neighborhood $U$ with $\overline{V} \subseteq U \subseteq L$	for which there exists $s' \in (s,\, t)$ such that the inclusion $U \cap f_{\leq s'} \to L \cap f_{\leq t}$ is $\HS_n$.
	We will show that $\Sigma_{s, t}$ has the following three properties:
	\begin{enumerate}
		 \item Any point $x \in L \cap f_{\leq s}$ has a neighborhood $V_x \in \Sigma_{s,t}$.
		 \item If $V_1,\, V_2 \in \Sigma_{s,t}$ then $V_1 \cup V_2 \in \Sigma_{s,t}$.
		 \item For each $V \in \Sigma_{s,t}$ the inclusion 
		 $V \cap f_{\leq s} \to L \cap f_{\leq t}$ is $\HS_n$.
	\end{enumerate}

@@

\begin{proof}[Proof of \cref{t:strong local connectedness implies q-tameness}]
	By definition, the sublevel set filtration of $f$ is q-tame if and only if the inclusion $f_{\leq s} \to f_{\leq t}$ is $\HS$ for all pairs $s < t$. 
	Choose $u \in \R$ with $u > t$ and let $g \colon Y \to \R$ be the restriction of $f$ to the sublevel set $Y=f_{\leq u}$. 
	Since we assume $f$ to induce a strongly $\HLC$ sublevel set filtration, by \cref{l:restriction of LHS filtration} the sublevel set filtration of $g$ is strongly $\HLC$ below $u$. 
	Clearly, the sublevel set filtration of $g$ is also compact, and its domain $Y$ is locally compact being a compact Hausdorff space by assumption.
	Thus, we can apply \cref{l:key lemma for q-tameness} to the filtration $g_{\leq \bullet}$ with $C = f_{\leq s}$ and $L = Y$ to obtain that the inclusion $f_{\leq s} = C \cap g_{\leq s} \to L \cap g_{\leq t} = f_{\leq t}$ is $\HS$.
\end{proof}

Combining \cref{t:hlc to strong hlc} with \cref{t:strong local connectedness implies q-tameness} also yields the following result for continuous functions.

\begin{cor} \label{c:q-tameness for continuous functions}
	If the sublevel set filtration of a continuous function $f \colon X \to \R$ is compact and $\HLC$, then it is also q-tame.
\end{cor}

### 2021-07-26 inequalities.tex
Improve code

In this section we prove that $q$-tame persistence modules satisfy a general version of Morse inequalities that specializes to the usual Morse inequalities in the smooth context, as well as to the version used by Morse and Tompkins to prove their Unstable Minimal Surface Theorem.
We deduce these general inequalities from the existence of persistence diagrams as reviewed in \cref{thm:q-tame modules have barcodes}.

Recall that for a Morse function $f$ on a closed smooth manifold $X$, the usual Morse inequalities state that for any non-negative integer $n$ the following holds:
\begin{equation} \label{equ:morse_inequalities}
\sum_{d=0}^n \ (-1)^{n-d} ( \# \crit_{d}(f) - \beta_{d}(X) ) \ \geq\  0,  
\end{equation}
where $\crit_{d}(f)$ is the set of critical points of $f$ with index $d$ and $\beta_{d}(X)$ is the $d\th$ Betti number of $X$.

@@

This approach of counting homological changes instead of critical points is also what Morse used in the non-smooth setting of functional topology.
To keep track of the number of $d$-dimensional homological events at filtration value $t$ that persist for at least time $\epsilon > 0$ but not indefinitely, \citet{Morse.1940} defined the $(d, t, \epsilon)$-\textit{cap numbers} of a filtration, which, when the associated persistence module has a barcode and is continuous from above, correspond to the number of finite bars with length greater than $\epsilon$ in the $d\th$ barcode with left endpoint $t$, plus the number of finite bars with length greater than $\epsilon$ in the $(d-1)\th$ barcode with right endpoint $t$ (killed by something $d$-dimensional).

Working entirely in the algebraic setting, we will now fix a graded q-tame persistence module $M$.
Of course, one may think of $M$ as the persistent homology of a q-tame filtration, but $M$ could for example also arise as the filtered Floer homology of some Hamiltonian on a symplectic manifold.
By \cref{thm:q-tame modules have barcodes}, $M$ has a persistence diagram in every degree $d$ with multiplicity function denoted $\mathfrak{m}_d \colon \multiplicityDomain \to \N$.
In analogy to Morse's definitions, we may define the $(d, t, \epsilon)$-\textit{cap number} of our graded q-tame persistence module $M$ in terms of its persistence diagram as
\begin{equation*}
m_{d}^{\epsilon}(t) =
\sum_{\substack{p \in \R \\ t - p > \epsilon}} \mathfrak{m}_{d-1}(p, t) +
\sum_{\substack{q \in \R \\ q - t > \epsilon}} \mathfrak{m}_d(t, q).
\end{equation*}
Note that finiteness of these quantities is ensured by the q-tameness of $M$ and the use of a non-zero $e$ bounding below the persistence of the considered features.
To see the necessity of this second condition, consider the q-tame persistence module given by the infinite product $\prod_{n \in \N_{> 0}} C([0,1/n))$.

Whenever the sums below are well-defined, we also consider the \emph{$(d,\epsilon)$-cap numbers} 
\begin{equation} \label{e:cap numbers}
m_{d}^{\epsilon} =
\sum_{t} m_{d}^{\epsilon}(t) \ =
\sum_{\substack{(p,q) \in \multiplicityDomain \\ q - p > \epsilon \\ q \neq \infty}} \mathfrak{m}_{d-1}(p,q) \ +
\sum_{\substack{(p,q) \in \multiplicityDomain \\ q - p > \epsilon \\ p \neq -\infty}}\mathfrak{m}_d(p,q).
\end{equation} 
They are well-defined, for example, if $M$ is the persistent homology of the sublevel set filtration of a bounded function.
This can be deduced from the following general result.

\begin{thm}
    Assume that $N$ is a q-tame persistence module that is initially and eventually constant, i.e., that there exist $t_0$ and $t_1$ with $N_s = N_{t_0}$ for all $s \leq t_0$ and $N_u = N_{t_1}$ for all $u \geq t_1$, and let $\mathfrak{n}$ be the multiplicity function of the persistence diagram of $N$.
    Then for each $\epsilon > 0$ we have 
    \begin{equation*}
    \sum_{ \substack{ (p,q) \in \multiplicityDomain \\ q-p > \epsilon } } \mathfrak{n} (p,q) < \infty.
    \end{equation*}
\end{thm}

\begin{proof}
    We split the sum whose finiteness we want to show in two parts
    \begin{equation*}
    \sum_{ \substack{ (p,q) \in \multiplicityDomain \\ q-p > \epsilon } } \mathfrak{n} (p,q)
    \ = \!\!
    \sum_{ \substack{ (p,q) \in \multiplicityDomain \setminus T \\ q-p > \epsilon } } \mathfrak{n} (p,q)
    \ +
    \sum_{ \substack{ (p,q) \in T \\ q-p > \epsilon } } \mathfrak{n} (p,q)
    \end{equation*}
    and show for each summand separately that it is finite.
    Here, $T$ denotes the triangle
    \begin{equation*}
    T = \{(p,q) \in \multiplicityDomain \mid t_0 \leq p < q \leq t_1\}.
    \end{equation*}

    For the first summand, observe that because $N$ is constant below $t_0$ and constant above $t_1$, we have $\mathfrak{n}(p,q) = 0$ whenever one of $-\infty < p < t_0$ or $q < t_0$, $p > t_1$ or $t_1 < q < \infty$ holds.
    This implies 
    \begin{equation*}
    \sum_{ \substack{ (p,q) \in \multiplicityDomain \setminus T \\ q-p > e } } \mathfrak{n} (p,q)
    \ = \!
    \sum_{t_0 < q < t_1} \mathfrak{n} (-\infty,q)
    \ + \!
    \sum_{t_0 < p < t_1} \mathfrak{n} (p, \infty)
    \end{equation*}
    which is clearly finite because $N$ is q-tame.

    For the second summand, note that  
    \begin{equation*}
    \sum_{ \substack{ (p,q) \in T \\ q-p > \epsilon } } \mathfrak{n} (p,q)
    \ \, \leq \!\!
    \sum_{(p,q) \in T^{\geq e}} \mathfrak{n} (p,q),
    \end{equation*}
    where $T^{\geq \epsilon}$ is the smaller triangle 
    \begin{equation*}
    T^{\epsilon} = \{(p,q) \in T \mid q-p \geq \epsilon\}.
    \end{equation*}
    Thus, in order to prove the theorem, it suffices to show that we have 
    \begin{equation*}
    \sum_{(p,q) \in T^{e}} \mathfrak{n} (p,q) 
    \ < \ 
    \infty.
    \end{equation*}
    To do this, we consider open quadrants 
    \begin{equation*}
    Q(x, y) = \{ (p, q) \in \R^2 \mid p < x \text{ and } y < q \}.
    \end{equation*}
    Covering the compact set $T^{\epsilon}$ by the open quadrants $Q \left(x, x + \frac{\epsilon}{2} \right)$ for $x \in \R$, we may choose a finite subcover given by, say, $x_1,\dots, x_n$.
	We obtain
    \begin{equation*}
    \sum_{(p,q) \in T^{\geq e}} \mathfrak{n} (p,q) 
    \ \leq \
    \sum_{i=1}^n \sum_{\substack{(p, q) \in \\ Q (x_i, x_i + \frac{\epsilon}{2})}} \mathfrak{n}(p,q).
    \end{equation*}
    Each of the sums $\sum_{(p,q) \in Q \left(x_i, x_i + \frac{\epsilon}{2} \right)} \mathfrak{n}(p,q)$ over the quadrants $Q \left(x_i, x_i + \frac{\epsilon}{2} \right)$ is finite since $N$ is q-tame (which is where the name q-tame or \emph{quadrant}-tame comes from, see \cite[Section 3.8]{Chazal.2016a}).
\end{proof}

Comparing to the usual Morse inequalities, the cap numbers in dimension $d$ act like the number of critical points with index $d$.
As an analogue to the Betti numbers of the manifold appearing in the usual Morse inequalities, Morse defines quantities $p_{d}$ referred to as \emph{essential dimensions}, which, under the same assumptions as before, can be expressed in the language of persistence diagrams as
\begin{equation} \label{e:essential dimension}
p_{d} \ = \!\! \sum_{p \in \R \cup \{-\infty\}} \mathfrak{m}_d(p,\infty),
\end{equation}
which is also the dimension of the colimit of the degree $d$ part of $M$.

\begin{thm}\label{t:inequalities}
	Let $M$ be a graded q-tame persistence module with cap numbers $m_{d}^{\epsilon}$ and essential dimensions $p_{d}$ finite for all $d$ and $\epsilon$.
	If $\mathfrak{m}_d(-\infty, p) = 0$ for all $p \in \R \cup \{\infty\}$ and all $d$, then we have Morse inequalities
    \begin{equation} \label{e:morse inequalities}
    \sum_{d=0}^n \ (-1)^{n-d} (m^{\epsilon}_{d} - p_{d}) \ \geq\  0
    \end{equation}
    for any dimension $n$ and any $\epsilon > 0$.
\end{thm}

\begin{proof}
    Using a telescopic sum argument, the claimed inequalities can be seen to be equivalent to the existence of a sequence $(\nu_d)_d$ of non-negative integers with $m^{\epsilon}_{d} - p_{d} = \nu_{d-1} + \nu_{d}$.
    Since we assume $\mathfrak{m}_d(-\infty, p) = 0$ for all $p$, we have
    \begin{align*}
    m^{\epsilon}_{d}
    \ &=
    \sum_{\substack{(p,q) \in \multiplicityDomain \\ q - p > \epsilon \\ q \neq \infty}} \mathfrak{m}_{d-1}(p,q)
    \ +
    \sum_{\substack{(p,q) \in \multiplicityDomain \\ q - p > \epsilon \\ p \neq -\infty}}\mathfrak{m}_d(p,q)
    \\
    &=
    \sum_{\substack{(p,q) \in \R^2 \\ q - p > \epsilon }} \mathfrak{m}_{d-1}(p,q)
    \ +
    \sum_{\substack{(p,q) \in \multiplicityDomain \\ q - p > \epsilon \\ p \neq -\infty}}\mathfrak{m}_d(p,q)
    \end{align*}
    and
    \begin{equation*}
    p_{d}
    \ = \!\!
    \sum_{p \in \R \cup \{-\infty\}} \mathfrak{m}_d(p,\infty)
    \ = \
    \sum_{p \in \R} \mathfrak{m}_d(p,\infty),
    \end{equation*}
    which yields
    \begin{align*}
    m^{\epsilon}_{d} - p_{d} \
    &= 
    \sum_{\substack{(p,q) \in \R^2 \\ q - p > \epsilon }} \mathfrak{m}_{d-1}(p,q)
    \ +
    \sum_{\substack{(p,q) \in \multiplicityDomain \\ q - p > \epsilon \\ p \neq -\infty}}\mathfrak{m}_d(p,q)
    \ - \
    \sum_{p \in \R} \mathfrak{m}_d(p,\infty)
    \\
    &=
    \sum_{\substack{(p,q) \in \R^2 \\ q - p > \epsilon }} \mathfrak{m}_{d-1}(p,q)
    \ + \!
    \sum_{\substack{(p,q) \in \R^2 \\ q - p > \epsilon }}\mathfrak{m}_d(p,q),
    \end{align*}
    so the sequence given by $\nu_d = \sum_{\substack{(p,q) \in \R^2 \\ q - p > \epsilon }}\mathfrak{m}_d(p,q)$ satisfies $m^{\epsilon}_{d} - p_{d} = \nu_{d-1} + \nu_{d}$ as desired.
\end{proof}

As we have remarked, the finiteness assumptions are satisfied if $M$ is initially and eventually constant, so as a special case, the theorem yields Morse inequalities for any bounded real-valued function whose sublevel set filtration has q-tame persistent homology, which includes classical Morse functions $f$ on closed smooth manifolds $X$.
As we have outlined while motivating the definition of the cap numbers, our inequalities as presented in \eqref{e:morse inequalities} agree with the classical inequalities \eqref{equ:morse_inequalities} because, in this setting, $\beta_d(X) = p_d$ and $m_d^{\epsilon} = \crit_d(f)$ for $\epsilon > 0$ smaller than the minimal difference between any two critical values of $f$.

>>>>

In this section we prove that $q$-tame persistence modules satisfy a general version of Morse inequalities that specializes to the usual Morse inequalities in the smooth context, as well as to the version used by Morse and Tompkins to prove their Unstable Minimal Surface Theorem.
We deduce these general inequalities from the existence of persistence diagrams as reviewed in \cref{t:q-tame modules have barcodes}.

Recall that for a Morse function $f$ on a closed smooth manifold $X$, the usual Morse inequalities state that for any non-negative integer $n$ the following holds:
\begin{equation} \label{e:morse_inequalities}
\sum_{d=0}^n \ (-1)^{n-d} \big( \# \crit_{d}(f) - \beta_{d}(X) \big) \ \geq \ 0,
\end{equation}
where $\crit_{d}(f)$ is the set of critical points of $f$ with index $d$ and $\beta_{d}(X)$ is the $d\th$ Betti number of $X$.

@@

This approach of counting homological changes instead of critical points is also what Morse used in the non-smooth setting of functional topology.
To keep track of the number of $d$-dimensional homological events at filtration value $t$ that persist for at least time $\epsilon > 0$ but not indefinitely, \citet{Morse.1940} defined the $(d, t, \epsilon)$-\emph{cap numbers} of a filtration, which, when the associated persistence module has a barcode and is continuous from above, correspond to the number of finite bars with length greater than $\epsilon$ in the $d\th$ barcode with left endpoint $t$, plus the number of finite bars with length greater than $\epsilon$ in the $(d-1)\th$ barcode with right endpoint $t$ (killed by something $d$-dimensional).

Working entirely in the algebraic setting, we will now fix a graded q-tame persistence module $M$.
Of course, one may think of $M$ as the persistent homology of a q-tame filtration, but $M$ could for example also arise as the filtered Floer homology of some Hamiltonian on a symplectic manifold.
By \cref{t:q-tame modules have barcodes}, $M$ has a persistence diagram in every degree $d$ with multiplicity function denoted $\m_d \colon \multiplicityDomain \to \N$.
In analogy to Morse's definitions, we may define the $(d, t, \epsilon)$-\emph{cap number} of our graded q-tame persistence module $M$ in terms of its persistence diagram as
\begin{equation*}
m_{d}^{\epsilon}(t) =
\sum_{\substack{p \in \R \\ t - p > \epsilon}} \m_{d-1}(p, t) +
\sum_{\substack{q \in \R \\ q - t > \epsilon}} \m_d(t, q).
\end{equation*}
Note that finiteness of these quantities is ensured by the q-tameness of $M$ and the use of a non-zero $e$ bounding below the persistence of the considered features.
To see the necessity of this second condition, consider the q-tame persistence module given by the infinite product $\prod_{n \in \N_{> 0}} C([0,1/n))$.

Whenever the sums below are well-defined, we also consider the \emph{$(d,\epsilon)$-cap numbers} 
\begin{equation} \label{e:cap numbers}
m_{d}^{\epsilon} =
\sum_{t} m_{d}^{\epsilon}(t) \ =
\sum_{\substack{(p,q) \in \multiplicityDomain \\ q - p > \epsilon \\ q \neq \infty}} \m_{d-1}(p,q) \ +
\sum_{\substack{(p,q) \in \multiplicityDomain \\ q - p > \epsilon \\ p \neq -\infty}}\m_d(p,q).
\end{equation} 
They are well-defined, for example, if $M$ is the persistent homology of the sublevel set filtration of a bounded function.
This can be deduced from the following general result.

\begin{thm}
	Assume that $N$ is a q-tame persistence module that is initially and eventually constant, i.e., that there exist $t_0$ and $t_1$ with $N_s = N_{t_0}$ for all $s \leq t_0$ and $N_u = N_{t_1}$ for all $u \geq t_1$, and let $\n$ be the multiplicity function of the persistence diagram of $N$.
	Then for each $\epsilon > 0$ we have 
	\begin{equation*}
	\sum_{ \substack{ (p,q) \in \multiplicityDomain \\ q-p > \epsilon } } \n (p,q) < \infty.
	\end{equation*}
\end{thm}

\begin{proof}
	We split the sum whose finiteness we want to show in two parts
	\begin{equation*}
	\sum_{ \substack{ (p,q) \in \multiplicityDomain \\ q-p > \epsilon } } \n (p,q)
	\ = \!\!
	\sum_{ \substack{ (p,q) \in \multiplicityDomain \setminus T \\ q-p > \epsilon } } \n (p,q)
	\ +
	\sum_{ \substack{ (p,q) \in T \\ q-p > \epsilon } } \n (p,q)
	\end{equation*}
	and show for each summand separately that it is finite.
	Here, $T$ denotes the triangle
	\begin{equation*}
	T = \{(p,q) \in \multiplicityDomain \mid t_0 \leq p < q \leq t_1\}.
	\end{equation*}

	For the first summand, observe that because $N$ is constant below $t_0$ and constant above $t_1$, we have $\n(p,q) = 0$ whenever one of $-\infty < p < t_0$ or $q < t_0$, $p > t_1$ or $t_1 < q < \infty$ holds.
	This implies 
	\begin{equation*}
	\sum_{ \substack{ (p,q) \in \multiplicityDomain \setminus T \\ q-p > e } } \n (p,q)
	\ = \!
	\sum_{t_0 < q < t_1} \n (-\infty,q)
	\ + \!
	\sum_{t_0 < p < t_1} \n (p, \infty)
	\end{equation*}
	which is clearly finite because $N$ is q-tame.

	For the second summand, note that
	\begin{equation*}
	\sum_{ \substack{ (p,q) \in T \\ q-p > \epsilon } } \n (p,q)
	\ \, \leq \!\!
	\sum_{(p,q) \in T^{\geq e}} \n (p,q),
	\end{equation*}
	where $T^{\geq \epsilon}$ is the smaller triangle 
	\begin{equation*}
	T^{\epsilon} = \{(p,q) \in T \mid q-p \geq \epsilon\}.
	\end{equation*}
	Thus, in order to prove the theorem, it suffices to show that we have 
	\begin{equation*}
	\sum_{(p,q) \in T^{e}} \n (p,q) 
	\ < \ 
	\infty.
	\end{equation*}
	To do this, we consider open quadrants 
	\begin{equation*}
	Q(x, y) = \{ (p, q) \in \R^2 \mid p < x \text{ and } y < q \}.
	\end{equation*}
	Covering the compact set $T^{\epsilon}$ by the open quadrants $Q \left(x, x + \frac{\epsilon}{2} \right)$ for $x \in \R$, we may choose a finite subcover given by, say, $x_1,\dots, x_n$.
	We obtain
	\begin{equation*}
	\sum_{(p,q) \in T^{\geq e}} \n (p,q) 
	\ \leq \
	\sum_{i=1}^n \sum_{\substack{(p, q) \in \\ Q (x_i, x_i + \frac{\epsilon}{2})}} \n(p,q).
	\end{equation*}
	Each of the sums $\sum_{(p,q) \in Q \left(x_i, x_i + \frac{\epsilon}{2} \right)} \n(p,q)$ over the quadrants $Q \left(x_i, x_i + \frac{\epsilon}{2} \right)$ is finite since $N$ is q-tame (which is where the name q-tame or \emph{quadrant}-tame comes from, see \cite[Section 3.8]{Chazal.2016a}).
\end{proof}

Comparing to the usual Morse inequalities, the cap numbers in dimension $d$ act like the number of critical points with index $d$.
As an analogue to the Betti numbers of the manifold appearing in the usual Morse inequalities, Morse defines quantities $p_{d}$ referred to as \emph{essential dimensions}, which, under the same assumptions as before, can be expressed in the language of persistence diagrams as
\begin{equation} \label{e:essential dimension}
p_{d} \ = \!\! \sum_{p \in \R \cup \{-\infty\}} \m_d(p,\infty),
\end{equation}
which is also the dimension of the colimit of the degree $d$ part of $M$.

\begin{thm} \label{t:inequalities}
	Let $M$ be a graded q-tame persistence module with cap numbers $m_{d}^{\epsilon}$ and essential dimensions $p_{d}$ finite for all $d$ and $\epsilon$.
	If $\m_d(-\infty, p) = 0$ for all $p \in \R \cup \{\infty\}$ and all $d$, then we have Morse inequalities
	\begin{equation} \label{e:morse inequalities}
	\sum_{d=0}^n \ (-1)^{n-d} (m^{\epsilon}_{d} - p_{d}) \ \geq\ 0
	\end{equation}
	for any dimension $n$ and any $\epsilon > 0$.
\end{thm}

\begin{proof}
	Using a telescopic sum argument, the claimed inequalities can be seen to be equivalent to the existence of a sequence $(\nu_d)_d$ of non-negative integers with $m^{\epsilon}_{d} - p_{d} = \nu_{d-1} + \nu_{d}$.
	Since we assume $\m_d(-\infty, p) = 0$ for all $p$, we have
	\begin{align*}
	m^{\epsilon}_{d}\ &=
	\sum_{\substack{(p,q) \in \multiplicityDomain \\ q-p > \epsilon \\ q \neq \infty}} \m_{d-1}(p,q) \ +
	\sum_{\substack{(p,q) \in \multiplicityDomain \\ q-p > \epsilon \\ p \neq -\infty}} \m_d(p,q) \\ &=
	\sum_{\substack{(p,q) \in \R^2 \\ q-p > \epsilon }} \m_{d-1}(p,q) \ +
	\sum_{\substack{(p,q) \in \multiplicityDomain \\ q-p > \epsilon \\ p \neq -\infty}} \m_d(p,q)
	\end{align*}
	and
	\begin{equation*}
	p_{d} \ = \!\!
	\sum_{p \in \R \cup \{-\infty\}} \m_d(p,\infty) \ = \
	\sum_{p \in \R} \m_d(p,\infty),
	\end{equation*}
	which yields
	\begin{align*}
	m^{\epsilon}_{d} - p_{d} \ &= 
	\sum_{\substack{(p,q) \in \R^2 \\ q-p > \epsilon}} \m_{d-1}(p,q) \ +
	\sum_{\substack{(p,q) \in \multiplicityDomain \\ q-p > \epsilon \\ p \neq -\infty}} \m_d(p,q)	\ - \
	\sum_{p \in \R} \m_d(p,\infty) \\	&=
	\sum_{\substack{(p,q) \in \R^2 \\ q-p > \epsilon}} \m_{d-1}(p,q)
	\ + \!
	\sum_{\substack{(p,q) \in \R^2 \\ q-p > \epsilon}} \m_d(p,q),
	\end{align*}
	so the sequence given by $\nu_d = \sum_{\substack{(p,q) \in \R^2 \\ q-p > \epsilon}} \m_d(p,q)$ satisfies $m^{\epsilon}_{d} - p_{d} = \nu_{d-1} + \nu_{d}$ as desired.
\end{proof}

As we have remarked, the finiteness assumptions are satisfied if $M$ is initially and eventually constant, so as a special case, the theorem yields Morse inequalities for any bounded real-valued function whose sublevel set filtration has q-tame persistent homology, which includes classical Morse functions $f$ on closed smooth manifolds $X$.
As we have outlined while motivating the definition of the cap numbers, our inequalities as presented in \eqref{e:morse inequalities} agree with the classical inequalities \eqref{e:morse_inequalities} because, in this setting, $\beta_d(X) = p_d$ and $m_d^{\epsilon} = \crit_d(f)$ for $\epsilon > 0$ smaller than the minimal difference between any two critical values of $f$.
